Any way the brain blows? The nature of decision noise in random exploration

### 2022-12-12 manuscript/Older drafts/Backup2/Internal-External Noise/Internal-External Noise.tex
updates

\title{What is the nature of decision noise in random exploration?}

@@

	\newpage
	\begin{abstract}
	[150 words for Nature Human Behavior, currently 162]
	Human decision making appears to contains components that are inherently random. While this behavioral variability has traditionally been seen as noise, recent work suggests that random choices may actually be adaptive. An example arises when deciding between exploring unknown options or exploiting options we know well. Theory shows that a little randomness in explore-exploit decisions is remarkably effective, leading to greater rewards overall. Meanwhile, experiments show that people use such `random exploration' in practice,  increasing their behavioral variability when it is more valuable to explore. Despite this progress, the nature of adaptive decision noise is unknown and it is unclear whether it is generated internally, from stochastic processes in the brain, or externally, from stochastic stimuli in the world? Here we show that, while both types of noise are present in explore-exploit decisions, random exploration is dominated by internal noise. This suggests that random exploration depends on adaptive noise processes in the brain which are subject to (perhaps unconscious) cognitive control.

@@

	Imagine trying to decide where to go to dinner with a friend. You can go to your favorite restaurant that you both really enjoy and always go to, or you can try the new restaurant that just opened and about which you know nothing. Such decisions, in which we must choose between a well-known exploit option and a lesser known explore option, are known as explore-exploit decisions.  From a theoretical perspective, making optimal explore-exploit choices, i.e. choices that maximize long-term reward, is incredibly hard \citep{eegittins74, eeBasu18}. In part because of this difficulty, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \citep{, eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}.

	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90}. In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise', meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance. In theory, such `random exploration,' is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11} ADDED TWO NEW PAPERS ON THOMPSON SAMPLING, NOT SURE WHETHER THEY ARE OKAY.

	Recently we have shown that humans appear to actually use random exploration and actively adapt their decision noise to solve simple explore-exploit problems \citep{wilson2014}.  The key manipulation in the task is the horizon condition, i.e. the number of decisions remaining to make. Increasing the horizon makes exploration more valuable as there is more time to use the information gained by exploration to maximize future rewards. For example, if you are leaving town tomorrow (short horizon), you will probably exploit the restaurant you know and love, but if you are in town for a while (long horizon), you would be more likely to explore the new place. Using such a horizon manipulation we found that people have greater decision noise in the long versus the short horizon condition.

	However, a key limitation of this work was that the source of the decision noise used for exploration is unknown. Of particular interest is whether the adaptive decision noise that is linked to exploration arises externally, in the input from the world, or is generated internally, within the brain. In the restaurant case, an example of external noise would be if you happen to spot an old friend walking in to one of the restaurants. Conversely, internal noise would arise from stochastic neural processes tossing a metaphorical coin in your head. Previous work makes a strong case for both types of noise being relevant to behavior. For instance, external, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than internal noise \citep{eeBrunton13}. Conversely internal, neural noise is thought to drive exploratory singing behavior in song birds  \citep{songbird2} and the generation and control of this internal noise has been linked to specific neural structures \citep{songbird2}.

	In this paper, we investigate which source of noise, internal vs external, drives random exploration in humans in a simple explore-exploit task adapted from our previous work \citep{wilson2014}. To distinguish between the two types of noise, we had people make the exact same explore-exploit decision twice. If decision noise is purely externally driven, then people's choices should be identical both times, that is their choices should be consistent, since the stimulus is the same both times. Meanwhile, if noise is internally driven their choices should be inconsistent, since the internal noise is different both times. By analyzing behavior on this task in both a model-free and model-based manner, we show that, while both types of noise are present in explore-exploit decisions, the contribution of internal noise to random exploration far exceeds that contributed by the stimulus.

@@

	Crucially, before people make their first choice in the Horizon Task, they are given information about the mean payoff from each bandit, in the form of four example plays distributed either unequally between bandits (i.e. 1 play of one bandit, 3 plays of the other) or equally (2 plays each). These example plays allow us to manipulate exactly what people know about each option before they make their first choice. Thus, by giving people the exact same example plays twice in two separate games (separated by several minutes in time so as to avoid detection), the example plays allow us to probe how participants respond to the exact same explore-exploit choice twice.

@@

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfiga.pdf}
			\caption{A key additional manipulation here is repeated games. Each pair of repeated games with identical example trials will appear twice during the experiment. We setup the repeated games such that they are at least 5 games apart from each other. A model-free measure of choice inconsistency which reflex the underlying decision noise is defined as the proportion of inconsistent choices for repeated games.}
			\label{fig:taskfig}
		\end{center}
	\end{figure}

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/line_modelfree.png}
			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions.}
			\label{fig:modelfree}
		\end{center}
	\end{figure}

@@

	To quantify choice inconsistency we computed the frequency with which participants made different responses for pairs of repeated games (Figure \ref{fig:mf2}). Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions (STATS), suggesting that not all of the noise was stimulus driven. In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (F(1, 236) = 37.98, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is internal.

@@

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/theory_da_info.png}
			\caption[Both external and internal noise contribute to the choice variability in random exploration]{Both external and internal noise contribute to the choice variability in random exploration. For both [1 3] and [2 2] condition, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only external noise exists where people should behave identically in repeated games. Also, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only internal noise exists where people treat repeated games independently. }
			\label{fig:mf2}
		\end{center}
	\end{figure}

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/hyperprior.png}
			\caption[Posterior distributions over the group-level noise standard deviation of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise standard deviation. Both internal and external noises are nonzero, and internal noise has a much greater magnitude. Posterior distributions over the group-level means of the change of external and internal noise standard deviation. Internal noise increases significantly with horizon, external noise increases with horizon as well.}
			%COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'
			\label{fig:mb1}
		\end{center}
	\end{figure}

@@

	\section*{Discussion}
	In this paper, we investigated whether random exploration is driven by internal noise, putatively arising in the brain, or external noise, arising from the environment. Using a version of the Horizon Task with repeated games, we found that evidence for both types of noise in explore-exploit decisions. In addition, we see that both internal and external noise increase with horizon, but that the horizon effect is much larger for internal noise.  Taken together our results suggest that random exploration, i.e. the use and adaptation of decision noise to drive exploration, is driven primarily by internal noise.

@@

	Taken at face value, the horizon-dependent increase in internal noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain. This is in line with work in the bird song literature in which song variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}. In addition, this work is consistent with a recent report from \cite{ebitz17} in which the behavioral variability of monkeys in an `explore' state was also tied to internal rather than external sources of noise.

	Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating internal levels of noise. Indeed, manipulation of the NE system has been found to change behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.  In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}

@@

	80 participants (ages 18-25, 43 male, 37 female) from the University of Arizona undergraduate subject pool participated in the experiment. 20 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}. This left 60 for the main analysis. Note that including the 20 badly performing subjects did not change the main results (Supplementary Figures \ref{fig:modelfree2}-\ref{fig:mb12})% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figall/line_modelfree.png}
			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions.}
			\label{fig:modelfree2}
		\end{center}
	\end{figure}

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figall/theory_da_info.png}
			\caption[Both external and internal noise contribute to the choice variability in random exploration]{Both external and internal noise contribute to the choice variability in random exploration. For both [1 3] and [2 2] condition, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only external noise exists where people should behave identically in repeated games. Also, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only internal noise exists where people treat repeated games independently. }
			\label{fig:mf22}
		\end{center}
	\end{figure}
	\newpage
	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figall/hyperprior.png}
			\caption[Posterior distributions over the group-level noise standard deviation of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise standard deviation. Both internal and external noises are nonzero, and internal noise has a much greater magnitude. Posterior distributions over the group-level means of the change of external and internal noise standard deviation. Internal noise increases significantly with horizon, external noise increases with horizon as well.}
			%COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'
			\label{fig:mb12}
		\end{center}
	\end{figure}
	%No differences of task-related behavior (accuracy, directed-exploration, random-exploration and choice inconsistency) were found between ages, genders, races and ethnicities. (\ref{appendix:participants})

@@

	The number of games participants played depended on how well they performed, which acted as the primary incentive for performing the task. Thus, the better participants performed, the sooner they got to leave the experiment. On average, participants played 151.6 games (minimum = 90 games, maximum = 192 games) and the whole task lasted between 12.34 and 32.74 minutes (mean 23.31 minutes).%  IS THIS NUMBER CORRECT?  I THOUGHT IT WAS MORE LIKE AN HOUR?%(\ref{appendixb})
	THIS IS ONLY THE TASK TIME, NOT INCLUDING BASELINE PUPIL OR THE TWO CURIOSITY SURVEY OR RA SETUP TIME

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfigb.pdf}
			\caption{Time line of a horizon 6 game.  The game begins with four forced choice trials in which participants are instructed which option to play.   After these forced trials they are free to choose between the two options for the rest of the game (6 trials in this case).
			%COMBINE THIS WITH FIGURE 7 FOR TASK CONDITIONS
			%CAN YOU MAKE THESE FIGURES LOOK NICER I.E. MORE LIKE MINE?  REDUCE LINE WIDTHS ETC ... IF NOT SEND ME POWER POINTS AND I WILL MAKE INTO PDFS}
			I WILL SEND YOU THE POWERPOINTS TOO}
			\label{fig:taskfigb}
		\end{center}
	\end{figure}

	The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose. We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice. The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice. After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions). (See Figure \ref{fig:taskfigb})

%	\begin{figure}[H]
%		\begin{center}
%			\includegraphics[width=\textwidth]{figures/taskfig2.PNG}
%			\caption{Task conditions}
%			\label{fig:taskfig2}
%		\end{center}
%	\end{figure}

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/Siyu_EEHorizon_2sigma_final.pdf}
			\caption{Hierarchical Bayesian model}
			\label{fig:model}
		\end{center}
	\end{figure}

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/precover.png}
			\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) %for horizon 1 games
			%COMBINE THESE TWO, HAVE TWO COLUMNS, LEFT IS HORIZON 1, RIGHT IS HORIZON 6.  REDUCE NUMBER OF SIGNIFICANT FIGURES IN p VALUES.  DON'T USE LEGEND, USE TEXT FUNCTION TO PUT TEXT ON PLOTS.  RED LINES CHANGE TO DOTTED BLACK LINES.  
			%XLABELS - SIMULATED $A$, YLABELS FIT $A$, I.E. MAKE SPECIFIC FOR EACH PARAMETER
			%USE ADDABCS FUNCTION TO ADD ABCs!
			}
			\label{fig:pararecover1}
		\end{center}
	\end{figure}

	THIS PLOT CHANGES EVERY TIME I RUN IT. THE $R^2$ VALUE FOR INFO BONUS AND INTERNAL NOISE ARE SIMILAR ALL THE TIME, BUT FOR EXTERNAL NOISE IT RANGES FROM 0.2 TO 0.7, FOR BIAS, FROM 0 TO 0.3
	%\begin{figure}[H]-
	%	\begin{center}
	%		\includegraphics[width=1\textwidth]{figures/paramrecover_2.png}
	%		\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) for horizon 6 games}
	%		\label{fig:pararecover2}
	%	\end{center}
	%\end{figure}
	%References
	%Plummer, M. (2003, March). JAGS: A program for analysis of Bayesian graphical models using Gibbs sampling. In Proceedings of the 3rd international workshop on distributed statistical computing (Vol. 124, p. 125).
	%Rescorla, R. A., & Wagner, A. R. (1972). A theory of Pavlovian conditioning: Variations in the effectiveness of reinforcement and nonreinforcement. Classical conditioning II: Current research and theory, 2, 64-99.

>>>>

\title{Any way the brain blows?  The nature of decision noise in random exploration}

@@

	\newpage
	\begin{abstract}

Human decision making is inherently variable. While this variability is often seen as a sign of suboptimality in human behavior, recent work suggests that randomness can actually be adaptive. An example  arises when we must choose between exploring unknown options or exploiting options we know well. A little randomness in these `explore-exploit' decisions is remarkably effective as it encourages us to explore options we might otherwise ignore. Moreover, people actually use such `random exploration' in practice,  increasing their behavioral variability when it is more valuable to explore. Despite this progress, the nature of adaptive `decision noise’ for exploration is unknown -- specifically whether it is generated internally, from stochastic processes in the brain, or externally, from stochastic stimuli in the world. Here we show that, while both internal and external noise drive variability in behavior, the noise driving random exploration is predominantly internal. This suggests that random exploration depends on adaptive noise processes in the brain which are subject to cognitive control.

%Human decision making is inherently variable. While this variability was originally seen as suboptimal, recent work suggests that randomness can actually be adaptive. An example arises when we must between exploring unknown options or exploiting options we know well. A little randomness in these `explore-exploit' decisions is remarkably effective, as it encourages to explore options we might otherwise ignore. Moreover, people actually use such `random exploration' in practice,  increasing their behavioral variability when it is more valuable to explore. Despite this progress, it is unknown whether variability for random exploration is generated externally, from stochastic stimuli in the world, or internally, from stochastic processes in the brain. That is do we explore by going any way the wind blows or any way our brains blow? Here we show that, while both types of noise are present that it is noise in our brains that dominates random exploration. This suggests that decision noise for exploration is subject to cognitive control.

@@

	Imagine trying to decide where to go to dinner with a friend. You can go to your favorite restaurant, the one you both really enjoy and always go to, or you can try the new restaurant that you know nothing about. Such decisions, in which we must choose between a well-known `exploit' option and a lesser known `explore' option, are known as explore-exploit decisions.  From a theoretical perspective, making optimal explore-exploit choices, i.e. choices that maximize long-term reward, is computationally intractable in most cases \citep{eegittins74, eeBasu18}. In part because of this difficulty, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \citep{, eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}.

	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90}. In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise,' meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance. In theory, such `random exploration,' is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11}.

	Using a simple gambling task, known as the Horizon Task, we have recently shown that humans appear to actually use random exploration and actively adapt their decision noise to solve simple explore-exploit problems \citep{wilson2014}.  The key manipulation in the Horizon Task is the horizon condition, i.e. the number of decisions remaining for the participant to make. Increasing the horizon makes exploration more valuable as there is more time to use the information gained by exploration to maximize future rewards. For example, if you are leaving town tomorrow (short horizon), you will probably exploit the restaurant you know and love, but if you are in town for a while (long horizon), you would be more likely to explore the new restaurant. Using such a horizon manipulation we found that people's behavior is more variable in long horizons than short horizons, suggesting that they use adaptive decision noise to solve the explore-exploit dilemma.

	A key limitation of our previous work was that the source of the decision noise used for exploration was left unknown.  In particular, due to the design of the task, we could not determine whether the adaptive decision noise that is linked to exploration arises externally, in the input from the world, or is generated internally, within the brain. 
	%internally, from random processes in the brain, or externally, in stochastic input from the world. In the restaurant case, internal noise would arise from stochastic neural processes tossing a metaphorical coin in your head.
	%$Of these, internal noise is the simplest to 
	%externally, in the input from the world, or is generated internally, within the brain. 
	In the restaurant case, an example of external noise would be if you happen to spot an old friend walking in to one of the restaurants. Such an event is random, in that you couldn't predict it, and has the potential to alter your behavior as you're likely to follow your friend in. Crucially, however, this `randomness' is very much in the world and if you saw the same friend go into the same restaurant at a later date you might follow them  again. 
	%If random exploration is driven by external noise, this implies that people adapt their susceptibility to random external with horizon. Thus
	%Exploration driven by external noise, 
	Conversely, internal noise would arise from stochastic neural processes tossing a metaphorical coin in your head. Such a process would not be influenced by the friend going into the restaurant, and if you saw the same friend again, you might make a different choice.

	Previous work makes a strong case for both external and internal noise being relevant to behavior. For instance, external, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than internal noise \citep{eeBrunton13}. Conversely internal, neural noise is thought to drive exploratory singing behavior in song birds  \citep{songbird2} and the generation and control of this internal noise has been linked to specific neural structures \citep{songbird2}.

	In this paper, we investigate which source of noise, external vs internal, drives random exploration in humans in a modified version of the Horizon Task. To distinguish between the two types of noise, we had people make the exact same explore-exploit decision twice. If decision noise is purely external, then people's choices should be identical both times, that is their choices should be consistent, since the stimulus is the same both times. Meanwhile, if noise is internally driven their choices should be inconsistent, since the internal noise is different both times. By analyzing behavior on this task in both a model-free and model-based manner, we show that, while both types of noise are present in explore-exploit decisions, the variability related to random exploration is dominated by internal noise.

@@

	Crucially, before people make their first choice in the Horizon Task, they are given information about the mean payoff from each bandit in the form of four example plays distributed either unequally between bandits (i.e. 1 play of one bandit, 3 plays of the other, the [1 3] condition) or equally (2 plays each, the [2 2] condition). These example plays allow us to manipulate exactly what people know about each option before they make their first choice. Thus, by giving people the exact same example plays twice in two separate games (separated by several minutes in time so as to avoid detection), the example plays allow us to probe how participants respond to the exact same explore-exploit choice twice.

@@

	\begin{figure}[hp]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfiga.pdf}
			\caption{BOB: ADD A PANEL A ABOVE THIS PANEL, SHOWING THE DYNAMICS FOR ONE GAME - LIKE IN HASHEM'S PAPER.  
			Schematic of the experiment. (A) Dynamics of an example horizon 6 game.  Here the first four trials are forced trials in which participants are instructed which option to play.  After the forced trials, participants are free to choose between the two options for the remainder of the game.  (B) Different possible states of the game after the first free choice over the course of the experiment. Overall participants play 160 such games, with varying horizon (1 vs 6), uncertainty condition ([1 3] vs [2 2]) and observed rewards.  In addition, all games are repeated (as Game 18 and 100 are here) such that participants will be faced with the exact same pattern of forced trials and exact same outcomes from those forced trials twice within each experiment.  These repeated games allow us to compute the relative contribution of external and internal noise by analyzing the extent to which choices are {\em consistent} across the repeated games.}%BOB: UP THE FONT SIZE FOR TEXT TO MATCH FONT SIZE FOR NUMBERS}
			\label{fig:taskfig}
		\end{center}
	\end{figure}

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/line_modelfree.png}
			\caption{BOB: DO THE YTICKS ON THE 0.1S NOT 0.15S!  MORE WHITE SPACE BETWEEN PANELS A AND B.
			Replication of previous findings. Both  $p(\mbox{low mean})$ (A) and $p(\mbox{high info})$ (B) increase with horizon suggesting that people use both random and directed exploration in this task.  }%BOB: SWITCH A AND B TO HAVE RANDOM FIRST}
			\label{fig:modelfree}
		\end{center}
	\end{figure}

@@

	To quantify choice inconsistency we computed the frequency with which participants made different responses for pairs of repeated games (Figure \ref{fig:mf2}). Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions (For [1 3] condition, t(59) = 11.78, p $<$ 0.001 for horizon 1, t(59) = 15.41, p $<$ 0.001 for horizon 6; For [2 2] condition, t(59) = 8.76, p $<$ 0.001 for horizon 1, t(59) = 17.02, p $<$ 0.001 for horizon 6.), suggesting that not all of the noise was stimulus driven. In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (F(1, 236) = 37.98 BOB DID YOU FIX THIS? DIDNT USE REPEATED ANOVA, WILL CHANGE, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is internal.

@@

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/theory_da_info.png}
			\caption{Model-free analysis suggests that both external and internal noise contribute to the choice variability in random exploration. For both the [1 3] (A) and [2 2] (B) condition, people show greater choice inconsistency in horizon 6 than horizon 1. However, the extent to which their choices are inconsistent lies between what is predicted by purely external and internal noise, suggesting that both noise sources influence the decision.}
			\label{fig:mf2}
		\end{center}
	\end{figure}

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/hyperprior.png}
			\caption{Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  internal and external noise. Both internal (A, B) and external (C,D) noises are nonzero (A, C) and change with horizon (B, D).  However, internal noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than external noise.}
			%COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'
			\label{fig:mb1}
		\end{center}
	\end{figure}

@@

	\section*{Discussion}
	In this paper, we investigated whether random exploration is driven by internal noise, putatively arising in the brain, or external noise, arising from the environment. Using a version of the Horizon Task with repeated games, we found evidence for both types of noise in explore-exploit decisions. In addition, we see that both internal and external noise increase with horizon, but that the horizon effect is much larger for internal noise.  Taken together our results suggest that random exploration, i.e. the use and adaptation of decision noise to drive exploration, is driven primarily by internal noise.

@@

	Taken at face value, the horizon-dependent increase in internal noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain. This is in line with work in the bird song literature in which variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}. In addition, this work is consistent with a recent report from \cite{ebitz17} in which the behavioral variability of monkeys in an `explore' state was also tied to internal rather than external sources of noise.

	Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating internal levels of noise. Indeed, changes in the NE system have been associated with changes behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.  In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}

@@

	80 participants (ages 18-25, 37 male, 43 female) from the University of Arizona undergraduate subject pool participated in the experiment. 20 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}. This left 60 for the main analysis. Note that including the 20 badly performing subjects did not change the main results (Supplementary Figures 1 - 3)% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS

	%SIYU: SORRY, WHAT DO YOU MEAN BY POPING THESE INTO A SEPARATE DOCUMENT, DO YOU MEAN I START A NEW LATEX DOC AND PUT THE SUPPLEMENTARY GRAPHS THERE, DO I STILL KEEP THE FIGURES HERE
	%BOB: SUPPLEMENTARY FIGURES SHOULD START NUMBERING FROM 1.  SO Supplementary Figure 1 etc ... Pop these into a separate document for sharing.  To do refs just hard code numbers.

	%No differences of task-related behavior (accuracy, directed-exploration, random-exploration and choice inconsistency) were found between ages, genders, races and ethnicities. (\ref{appendix:participants})

@@

	The number of games participants played depended on how well they performed, which acted as the primary incentive for performing the task. Thus, the better participants performed, the sooner they got to leave the experiment. On average, participants played 151.6 games (minimum = 90 games, maximum = 192 games) and the whole task lasted between 12.34 and 32.74 minutes (mean 23.31 minutes).%  IS THIS NUMBER CORRECT?  I THOUGHT IT WAS MORE LIKE AN HOUR?%(\ref{appendixb})
	%THIS IS ONLY THE TASK TIME, NOT INCLUDING BASELINE PUPIL OR THE TWO CURIOSITY SURVEY OR RA SETUP TIME

@@

	The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose. We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice. The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice. After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions),  Figure \ref{fig:taskfigb}.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfigb.pdf}
			\caption{BOB: DONT NEED THIS FIGURE IF YOU UPDATE EARLIER TASKFIGURE. Time line of a horizon 6 game.  The game begins with four forced choice trials in which participants are instructed which option to play.   After these forced trials they are free to choose between the two options for the rest of the game (6 trials in this case).
			%BOB: THIS FONT SIZE IS GOOD - MAYBE A LITTLE BIGGER FOR TEXT
			%COMBINE THIS WITH FIGURE 7 FOR TASK CONDITIONS
			%CAN YOU MAKE THESE FIGURES LOOK NICER I.E. MORE LIKE MINE?  REDUCE LINE WIDTHS ETC ... IF NOT SEND ME POWER POINTS AND I WILL MAKE INTO PDFS}
			%I WILL SEND YOU THE POWERPOINTS TOO
			}
			\label{fig:taskfigb}
		\end{center}
	\end{figure}
%	\begin{figure}[H]
%		\begin{center}
%			\includegraphics[width=\textwidth]{figures/taskfig2.PNG}
%			\caption{Task conditions}
%			\label{fig:taskfig2}
%		\end{center}
%	\end{figure}

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/Siyu_EEHorizon_2sigma_final.pdf}
			\caption{Schematic of the hierarchical Bayesian model using notation of BOB CITE WAGNERMAKERS AND LEE.}
			\label{fig:model}
		\end{center}
	\end{figure}

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/precover.png}
			\caption{Parameter recovery over the subject-level means of information bonus, $A$, spatial bias, $b$, internal noise variance, $\sigma_{int}$, and external noise variance, $\sigma_{ext}$, for horizon 1 (left column) and horizon 6 (right column) games. %for horizon 1 games
			%COMBINE THESE TWO, HAVE TWO COLUMNS, LEFT IS HORIZON 1, RIGHT IS HORIZON 6.  REDUCE NUMBER OF SIGNIFICANT FIGURES IN p VALUES.  DON'T USE LEGEND, USE TEXT FUNCTION TO PUT TEXT ON PLOTS.  RED LINES CHANGE TO DOTTED BLACK LINES.  
			%XLABELS - SIMULATED $A$, YLABELS FIT $A$, I.E. MAKE SPECIFIC FOR EACH PARAMETER
			%USE ADDABCS FUNCTION TO ADD ABCs!
			%BOB: BIT MORE SPACE BETWEEN ROWS. OTHERWISE GREAT!
			}
			\label{fig:pararecover1}
		\end{center}
	\end{figure}

%	THIS PLOT CHANGES EVERY TIME I RUN IT. THE $R^2$ VALUE FOR INFO BONUS AND INTERNAL NOISE ARE SIMILAR ALL THE TIME, BUT FOR EXTERNAL NOISE IT RANGES FROM 0.2 TO 0.7, FOR BIAS, FROM 0 TO 0.3
%BOB: I THINK THIS IS OK.
	%\begin{figure}[H]-
	%	\begin{center}
	%		\includegraphics[width=1\textwidth]{figures/paramrecover_2.png}
	%		\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) for horizon 6 games}
	%		\label{fig:pararecover2}
	%	\end{center}
	%\end{figure}
	%References
	%Plummer, M. (2003, March). JAGS: A program for analysis of Bayesian graphical models using Gibbs sampling. In Proceedings of the 3rd international workshop on distributed statistical computing (Vol. 124, p. 125).
	%Rescorla, R. A., & Wagner, A. R. (1972). A theory of Pavlovian conditioning: Variations in the effectiveness of reinforcement and nonreinforcement. Classical conditioning II: Current research and theory, 2, 64-99.
